Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"

### 2018-06-11 design.tex

Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance) .   
Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges. 
By spending more time together, participants are more likely to overcome barriers of professional terminology.
However, events that are too long may lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and project work.

A hack week requires a flexible workspace environment with ample opportunity for re-configuration. 
Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
Workspaces must also accommodate interactive project work where small teams can gather and work together (for more details, see also the SM, Sections 4.1.3 and 4.1.4).    
%Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
%However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
Fortunately, many universities are experimenting with other types of spaces: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.

Another important design consideration is group size (see also SM, Section 4.1.4).
In a large group, chances for random participant exchanges may be reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations.
We have found groups with sizes between 50 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group, but we encourage organizers to experiment with group size.

@@

Hack week outcomes depend strongly on the interest and engagement of participants.
Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week (see also SM; Section 4.2.4). A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 4.1.7). 
Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants. In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).

>>>>

Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance).
Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges.
By spending more time together, participants are more likely to overcome barriers of professional terminology.
However, events that are too long may also lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and project work.

Another important design consideration is the space. A hack week requires a flexible workspace environment with ample opportunity for re-configuration.
Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops.
Workspaces must also accommodate interactive project work where small teams can gather and work together (for more details, see also the SM, Sections 4.1.3 and 4.1.4).
%Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure.
%However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
Fortunately, many universities are experimenting with new types of spaces that allow these kinds of activities: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.

Another important design consideration is group size (see also SM, Section 4.1.4).
In a large group, chances for random participant exchanges may be reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations.
We have found groups with sizes between 40 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group, but we encourage organizers to experiment with group size.

@@

Hack week outcomes depend strongly on the interest and engagement of participants.
Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
The large variety in participant backgrounds and experiences--and the resulting range in personalities and objectives of attendees--requires careful, active facilitation of both taught and project components of a hack week (see also SM; Section 4.2.4). A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 4.1.7).
Community building is a core component of a hack week, and the design of hack weeks needs to include both very strong personalities and very shy participants. In particular, the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).

### 2018-06-11 themes.tex

To date, all hack weeks we have organized have been subject-specific, i.e.\ aimed at bringing together a community with a shared scientific interest, such as neuroscience.
Advantages to this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.
On the other hand, homogeneity may lead to \textit{group think} and inhibit new, creative solutions. 
In this case, it may be advantageous to design a hack week around a technique (e.g.\ Gaussian Processes) or modality (e.g.\ imaging), such as the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}}) meetings. 
For these events, building a shared vocabulary and understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.

>>>>

To date, all hack weeks we have organized have been subject-specific, i.e.\ aimed at bringing together a community with a shared scientific interest, such as neuroscience.
Advantages to this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.
On the other hand, homogeneity may lead to \textit{group think} and inhibit new, creative solutions.
In this case, it may be advantageous to design a hack week around a technique (e.g.\ Gaussian Processes) or modality (e.g.\ imaging), such as the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}}) meetings.
For these events, building a shared vocabulary and shared understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.

### 2018-06-11 what.tex

%Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. Additionally, they provide opportunities to build new collaborations and network. 
Our hack weeks combine structured, tutorial-style instruction with open-ended project work, providing opportunities for peer learning, networking, and building collaborations.  
In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).

The hackathon, a time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events. 
Hackathons originated from the open-source software movement and have historically focused on software and technology development. 
In recent years hackathons have evolved into a model providing opportunities for intensive, interdisciplinary collaboration \cite{Groen2015-cj} and education \cite{Kienzler2015-zu,Lamers2014-xf} in the sciences. 
Core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
Building on these core elements hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}. 
%``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
%``Contributive'' hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
%Finally, ``Communal'' hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.

@@

Hack weeks uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number and content of tutorials. This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities. They encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few. Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach and work aimed at improving the scientific community itself.

There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week might potentially not do any of them well. Because tutorials are not necessarily the major focus of a hack week, the knowledge gained by participants in these tutorials may be shallow. A hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require significant preparation from the side of participants. 
Because of these risks, organizers face a much larger degree of uncertainty, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, and facilitation of the large range of different types of activities at any given time (see also the SM, Section 4.2.4).

>>>>

%Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. Additionally, they provide opportunities to build new collaborations and network.
Our hack weeks combine structured, tutorial-style instruction with open-ended project work, providing opportunities for peer learning, networking, and for building collaborations.
In a space spanned by pedagogical focus as one dimension and focus on project work as the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).

The hackathon, a time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events.
Hackathons originated from the open-source software movement and have historically focused on software and technology development.
In recent years hackathons have evolved into a model providing opportunities for intensive, interdisciplinary collaboration \cite{Groen2015-cj} and education \cite{Kienzler2015-zu,Lamers2014-xf} in the sciences.
Core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
Building on these core elements, hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}.
%``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
%``Contributive'' hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
%Finally, ``Communal'' hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.

@@

Hack weeks uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number and content of tutorials. This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities. They encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few. Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: For example, our hack weeks explicitly encourage projects around outreach and work aimed at improving the scientific community itself.

There is, however, a major risk in the lack of focus: by wanting to do many things at once, a hack week might potentially not do any of them well. Because tutorials are not necessarily the major focus of a hack week, the knowledge gained by participants in these tutorials may be shallow. A hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require significant preparation from the side of participants.
Because of these risks, organizers face a much larger degree of uncertainty, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, and facilitation of the large range of different types of activities at any given time (see also the SM, Section 4.2.4).

### 2018-06-11 why.tex

\begin{itemize}
\item{\textit{Education and Training}:
While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning. This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}

@@

\item{\textit{Community Building}: Hack weeks provide a tremendous opportunity to catalyze community development through a shared interest in solving computational challenges with open source software. These events allow computationally minded researchers to break from the isolation of their academic departments and spark new collaborations.}

>>>>

\begin{itemize}
\item{\textit{Education and Training}:
While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning. This type of lateral knowledge transfer \cite{b:wilson-swc-lessons-2016} is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}

@@

\item{\textit{Community Building}: Hack weeks provide an opportunity to catalyze community development through a shared interest in solving computational challenges with open source software. These events allow computationally-minded researchers to break from the isolation of their academic departments and spark new collaborations.}
